Characterizing Signal Loss, Error, and Bias in the 21cm Reionization \\Power Spectrum: A Revised Study of PAPER-64

### 2018-04-25 PAPER.tex
Tue Apr 24 22:58:57 EDT 2018

We illustrate the way in which the quadratic estimator, Equation \ref{eq:QuadraticEstimator}, is intended to estimate the power spectrum of EoR in the presence of contamination by other terms by considering the simpler, but similar, case of estimating the variance of data.  We have specifically chosen a toy model where the data covariance is diagonal, as we expect the EoR signal to be.  We assume we have $N$ data points $\Delta_i$ which are the sum of a desired signal $\sigma_i$ and some contaminant $\phi_i$
\begin{equation}
\Delta_i = \sigma_i + \phi_i
\end{equation}
with 
\begin{equation}
\langle \sigma_i \rangle = 0; \; \langle \sigma^2_i \rangle = s; \; {\rm and}~\langle \sigma \sigma^T \rangle = s I_{N \times N} \equiv S
\end{equation}
and we wish to estimate $s$.  The contaminant has a similar structure for its covariance, and is assumed uncorrelated
\begin{equation}
\langle \sigma_i \rangle = 0; \; \langle \sigma^2_i \rangle = s; \; {\rm and}~\langle \sigma \sigma^T \rangle = s I_{N \times N} \equiv S
\end{equation}
$\langle \phi_i \rangle = 0$, 
$\langle \phi^2_i \rangle = g$, $\angle \phi \phi^T \rangle = g I_{N \times N} \equiv F$, 
$\langle \sigma_i \phi_j \rangle = 0$.
That is, we have the sum of $N$ independent samples of two random variables and we wish to estimate the variance of one of the random variables in the sum; in this case, $s$.
If we want to estimate $s$ using only the inverse covariance weighting, we find
\begin{equation}
\label{eq:CaseOneICW}
\hat{s} = \half {\cal F}^{-1} \left( \Delta^T \invC \PDeriv{C}{s}  \invC \Delta \right)
\end{equation}
and plugging in the relevant quantities gives
\begin{eqnarray}
\hat{s} & = & \half \frac{2}{N} (s+g)^2 \left( \Delta^T \frac{1}{s+g} I \frac{1}{s+g} \Delta \right) \\
& = & \frac{\Delta^T \Delta}{N}
\end{eqnarray}
We note that
\begin{equation}
\langle \hat{s} \rangle = s + g
\end{equation}
so in this case, {\it where the covariance structure of the contaminant is identical to the signal}, it is only possible to achieve an unbiased estimate by using some additional information to estimate the additive bias and subtract it.

If instead of the foreground contaminant being given by Eq \ref{eq:ToyOne}, suppose we now have
\begin{equation}
\Delta_i = \sigma_i + \phi m_i
\end{equation}
where again
\begin{eqnarray}
\langle \sigma_i \rangle & = & 0 \\
\langle \sigma^2_i \rangle & = & s \\
 \langle \sigma \sigma^T \rangle & = & s I_{N \times N} \equiv S
\end{eqnarray}
but now $\phi$ is a random variable and $m_i$ is a fixed function of $i$ with
\begin{eqnarray}
\langle \phi \rangle & = & 0 \\
\langle \phi^2 \rangle & = & g \\
 \langle \phi \phi^T \rangle & = & g m m^T \equiv G \\
 m^T m & = & 1
\end{eqnarray}
This is the case where each data point is an independent draw from a variable with a fixed variance $s$ we would like to estimate, and $m$ represents a mode which is correlated across many data points with correlation given by $m$, and amplitude in realization of $\phi$.
This is the essence of the toy model presented in Section \ref{sec:}

>>>>

We illustrate the way in which the quadratic estimator, Equation \ref{eq:QuadraticEstimator}, is intended to estimate the power spectrum of EoR in the presence of contamination by other terms by considering the simpler, but similar, case of estimating the variance of data.  We have specifically chosen a toy model where the data covariance is diagonal, as we expect the EoR signal to be.  We assume we have $N$ data points $\Delta_i$ which are the sum of a desired signal $\sigma_i$ and some contaminant $\phi_i$
\begin{equation}
\Delta_i = \sigma_i + \phi_i
\end{equation}
with 
\begin{equation}
\langle \sigma_i \rangle = 0; \; \langle \sigma^2_i \rangle = s; \; {\rm and}~\langle \sigma \sigma^T \rangle = s I_{N \times N} \equiv S
\end{equation}
and we wish to estimate $s$.  The contaminant in this case has a similar structure for its covariance, and is assumed uncorrelated with the signal
\begin{equation}
\langle \phi_i \rangle = 0; \; 
\langle \phi^2_i \rangle = f; \; 
\langle \phi \phi^T \rangle = f I_{N \times N} \equiv F; {\rm and}~ 
\langle \sigma_i \phi_j \rangle = 0
\end{equation}
%That is, we have the sum of $N$ independent samples of two random variables and we wish to estimate the variance of one of the random variables in the sum; in this case, $s$.
The estimator for $s$ using the inverse covariance weighting prescription is
\begin{equation}
\label{eq:CaseOneICW}
\hat{s} = \half {\cal F}^{-1} \left( \Delta^T \invC \PDeriv{C}{s}  \invC \Delta \right) 
= \frac{\Delta^T \Delta}{N}
\end{equation}
%and plugging in the relevant quantities gives
%\begin{eqnarray}
%\hat{s} & = & \half \frac{2}{N} (s+g)^2 \left( \Delta^T \frac{1}{s+g} I \frac{1}{s+g} \Delta \right) \\
%& = & \frac{\Delta^T \Delta}{N}
%\end{eqnarray}
We note that
\begin{equation}
\langle \hat{s} \rangle = s + g
\end{equation}
In this case, the quadratic estimator returns an entirely reasonable result, but, 
{\it since the covariance structure of the contaminant is identical to the signal}, it is only possible to achieve an unbiased estimate by using some additional information to estimate the additive bias and subtract it.

Now consider a case, chosen to be very similar to the toy model in \ref{sec:},  in which the data have an additive contaminant given by
\begin{equation}
\Delta_i = \sigma_i + \phi m_i
\end{equation}
where the properties of $\sigma_i$ are as before,
%\begin{eqnarray}
%\langle \sigma_i \rangle & = & 0 \\
%\langle \sigma^2_i \rangle & = & s \\
% \langle \sigma \sigma^T \rangle & = & s I_{N \times N} \equiv S
%\end{eqnarray}
but now $\phi$ is a random variable and $m_i$ is a fixed function of $i$ with
\begin{equation}
\langle \phi \rangle = 0; \;
\langle \phi^2 \rangle = f; \; 
 \langle \phi \phi^T \rangle = f m m^T \equiv F; \;
 m^T m = 1
\end{equation}
Here $m$ represents a mode which is correlated across many data points, with correlation given by $m m^T$, and, and the amplitude of its realization is $\phi$.
